---
title: "Prediction of Exercise Execution"
output: html_document
---
**Human Activity Recognition, Weight Lifting Exercise Dataset **

In contrast to studies to determine what activity a person is doing, this study evaluates participants execution of common weight lifting exercise, 4 done with incorrect form and one done per specification. This is identified as the classe, with "A"" being correct form and B through E being incorrect form mimicing common mistakes.
Data were collected via sensors at the bicep, wrist and hips of the participants.


**Load Libraries**
```{r}
#load libraries
library(caret)
library(randomForest)
library(corrplot)
```

**Load Data**
```{r}
train <- read.csv("pml-training.csv", na.strings=c("", "\"\"", "NA"))
test <- read.csv("pml-testing.csv", na.strings=c("", "\"\"", "NA"))
```

**First Look at the data**

There are 19622 obs. of  160 variables, but there are numerous columns that are entirely NA
```{r}
q<-sapply(train, function(x)all(!is.na(x)))
summary(q)
##alternately caret::nearZeroVar(train)
```


**Clean Data**

We will remove the NA columns from the train and test datasets, additionally, we will remove several factor variables.
```{r}
cleanTrain <- train[, colSums(!is.na(test)) != 0]
cleanTest <- test[, colSums(!is.na(test)) != 0]
cleanTrain <- subset (cleanTrain,select = -c(X,user_name,cvtd_timestamp, raw_timestamp_part_1, raw_timestamp_part_2, new_window, num_window))
cleanTest <-  subset (cleanTest,select = -c(X,user_name, cvtd_timestamp, raw_timestamp_part_1, raw_timestamp_part_2, new_window, num_window))
```
This leaves 19622 obs. of 53 variables, the last variable in the train dataset being the classe variable we are modeling a predction for.

**Exploratory Data Analysis**

A quick review of the distribution of the classe variable. Also, a plot of correlations to help in starting to identify variables of importance to the model. (the corrplot is poorly rendered in the HTML, but I included the code if someone wants to run in R Studio)
```{r}
plot(cleanTrain$classe,col=rainbow(5),main = "`classe` frequency plot")
m <-cor(cleanTrain[,-53])
#corrplot(m)
par(mfrow=c(1,3)) 
corrplot(m[,1:17])
corrplot(m[,18:35])
corrplot(m[,35:52])
```


**Partion Training Data**

Splitting the cleaned train dataset into a 70% training and 30% validation subset.
```{r}
set.seed=(4567)
partition <- createDataPartition(y = cleanTrain$classe, p = 0.7, list = FALSE)
trainingdata <- cleanTrain[partition, ]
validationdata  <- cleanTrain[-partition,]

```

**Create model from training data**

Using Random Forest in the caret package to model the training dataset.
```{r}
model <- train(classe ~ ., data = trainingdata, method = "rf", prox = TRUE,trControl = trainControl(method = "cv", number = 4, allowParallel = TRUE))

```

**Analysis**

A review of the variable importance in the model
```{r}
varImpPlot(model$finalModel,sort = TRUE, pch=19, main="Variable Importance")
```

**Validate Model**

Validating the model against the subsetted validation dataset.
```{r}
test_predict <- predict(model,validationdata)
confusionMatrix(test_predict,validationdata$classe)
```
The estimated accuracy of the model is 99.3% and the estimated out-of-sample error based on our fitted model applied to the cross validation dataset is 0.7%. 


**Predict Classe in Test Data**

Running the model against the test dataset.
```{r}
answers <-predict(model,cleanTest)
answers <- as.character(answers)
answers
```


**Data Citation**

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013. 

**Link to Data**

http://groupware.les.inf.puc-rio.br/static/WLE/WearableComputing_weight_lifting_exercises_biceps_curl_variations.csv